---
title: "Hierarchical Models, CLT and Delta Method"
author: " Your Name and id"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes    
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Hierarchical Models

Consider the following hierarchical model:

$$ T \sim \text{Gamma}(\alpha = 10, \beta = 4), \quad Y|T \sim \text{Pois}(T)$$
Write a function that models the random variable $Y$. This function must
take input values $\alpha$ and $\beta$ and output the value that the
random variable will take on a single run of the experiment.

```{r}
poisson_gamma <- function(alpha, beta){
  T <- rgamma(n = 1, shape = alpha, scale = beta)
  Y <- rpois(n = 1, lambda = T)
}
```

Use the replicate function to get 10000 values of the random variable
$Y$. Store these values in a variable 'y_empirical'.

```{r}
y_empirical <- replicate(10000,poisson_gamma(10,4))
```

Use the 'hist' function to calculate the empirical distribution of $Y$.

```{r}
hist(y_empirical)
```

Now get 10000 random values from the appropriate negative binomial
distribution, and store these in the variable 'y_theoretical'. Use the
'qqplot' function to compare the quantiles of 'y_empirical' to those of
'y_theoretical'. What do you notice?

```{r}
# k/p is the mean
y_theoretical <- replicate(10000,rnbinom(n = 80,size = 40, prob = .5)) 
par(mfrow = c(1,2))
hist(y_theoretical)
qqplot(y_theoretical,y_empirical)
```

# Limiting Distributions

## Maximum of Uniforms

Write a function 'uni_sample_max' that simulates the 'sample_max'
statistic on a random sample of size $n$ coming from the uniform
distribution on $(0, 1)$. Note that this function takes input $n$, the
sample size.

```{r}

```

Use the above function define the statistic:
$$ \hat{\theta} = n(1-X_{(n)}) $$ We know that this statistic converges
in distribution to $\text{Exp}(\lambda = 1)$. Call the function
'theta_hat' which must take input $n$, the sample size of the uniform
random sample.

```{r}

```

Write a function \`theta_hat_plot' which takes input $n$, and outputs
the qqplot of 10000 values coming from the replicates of 'theta_hat' and
10000 values coming from the $\text{Exp}(\lambda = 1)$ (use the 'rexp'
command for the latter).

```{r}

```

Evaluate 'theta_hat_plot' on values of $n$ in the vector
$(1, 2, 3, 5, 10, 20, 40, 50, 100, 1000)$.

```{r}

```

# Delta Method

## Odds for Binomial Distribution

Consider the estimator $\hat{\theta} = \frac{\hat{p}}{1-\hat{p}}$,
(where $\hat{p}$ is proportion of successes in $n$ Bernoulli trials) for
the parameter 'odds' $\theta = \frac{p}{1-p}$. Let
$g(p) = \frac{p}{1-p}$, then using the delta-method we know that
$$ \sqrt{n}(g(\hat(p)) - g(p)) \longrightarrow N(0, \sigma^2(g'(p))^2) $$

Write a function 'odds_hat' that takes input the sample size $n$, and
calculates the odds estimate for $n$ independent Bernoulli random
variable with $p=0.7$.

```{r}

```

Write a function \`odds_plot' which takes input $n$, and outputs the
qqplot of 10000 values coming from the replicates of 'odds_hat' and
10000 values coming from the $N(0, \sigma^2(g'(0.7))^2)$ (use the
'rnorm' command for the latter).

```{r}

```

Evaluate 'odds_plot' on values of $n$ in the vector
$(1, 2, 3, 5, 10, 20, 40, 50, 100, 1000)$.

```{r}

```

## Approximating $\frac{1}{\mu}$.

Consider sampling from $\text{Gamma}(\alpha =2, \beta = 4)$. Run a
simulation (like the one for 'odds') to show the following property of
the limiting distribution of the reciprocal of the sample mean

$$ \frac{\sqrt{n}\left(\frac{1}{\overline{X}} - \frac{1}{\mu}\right)}{\frac{1}{\overline{X}^2} S} \longrightarrow N(0,1)$$
where $\overline{X}$ is the sample mean, and $S = \sqrt{S^2}$.
